---
title: "Illustrating R Notebook Capabilities"
output:
  pdf_document: default
  html_notebook: default
---

R Notebook is interactive and more suited for exploratory analysis and immediate feedback. Both R Markdownare based on R Markdown, but R Notebook enhances the experience by allowing live code execution and interactive exploration.

In this notebook, we explore the marathon data from a CSV file. We will visualize the distribution of finish times and compare finish times for different runner types.

First, we load the data and necessary libraries.
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Load the data
marathon_data <- read.csv("Marathon.csv")

```
 
We then view the first 5 rows just to get a glimpse of what we are working with.
```{r}
# View the first few rows of the data
head(marathon_data)
```

Now, let's plot a histogram to examine the distribution of marathon finish times across all runners.
```{r}
# Basic histogram of Finish Time
ggplot(marathon_data, aes(x = Finish_Time)) + 
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black", alpha = 0.7) + 
  labs(title = "Distribution of Finish Times", x = "Finish Time (hours)", y = "Count")

```

Next, we will create a boxplot to compare the finish times across different runner types. This will help us understand the variations in performance for "Professional," "First-Timer," and "Frequent" runners.

```{r}

# Boxplot to compare Finish Time across different Types
ggplot(marathon_data, aes(x = Type, y = Finish_Time, fill = Type)) + 
  geom_boxplot() + 
  labs(title = "Finish Time by Type", x = "Runner Type", y = "Finish Time (hours)") + 
  theme_minimal()

```
  
  