---
title: "Interactive Graphics with Shiny Widgets"
output: 
  pdf_document: default
  html_notebook: default
runtime: shiny
---

## Introduction

In this example, we demonstrate how to use Shiny widgets to interactively explore marathon finish time data. You will be able to select a runner type and adjust the histogram bin width dynamically.

## Load Libraries and Data

First, we load the necessary libraries and the marathon data.

```{r}
# Load the necessary libraries
library(shiny)
library(ggplot2)
library(dplyr)

# Load the marathon data
marathon_data <- read.csv("Marathon.csv")
```

The first visualization is a histogram that displays the distribution of finish times for all runners. I implemented an interactive slider that allows the user to adjust the bin width of the histogram, which helps in examining the distribution at different levels.

```{r}
# Define UI for the application
ui <- fluidPage(
  
  # Title
  titlePanel("Interactive Marathon Finish Times"),
  
  # Dropdown to select the runner type
  selectInput("runner_type", "Select Runner Type:",
              choices = c("Professional", "First-Timer", "Frequents")),
  
  # Slider to adjust bin width for the histogram
  sliderInput("bin_width", "Select Bin Width:",
              min = 0.1, max = 1, value = 0.5, step = 0.1),
  
  # Plot output
  plotOutput("distPlot")
)

```

Next, we define the server logic to make sure the app responds to the user's selections. The key part here is the reactive expression that updates the data based on the selected runner type and the slider value for bin width.

```{r}
# Define server logic for the application
server <- function(input, output) {
  
  # Reactive expression to filter the data based on selected runner type
  filtered_data <- reactive({
    if (input$runner_type == "All") {
      marathon_data
    } else {
      filter(marathon_data, Type == input$runner_type)
    }
  })
  
  # Render the histogram dynamically based on the input
  output$distPlot <- renderPlot({
    data <- filtered_data()
    
    ggplot(data, aes(x = Finish_Time)) +
      geom_histogram(binwidth = input$bin_width, fill = "skyblue", color = "black", alpha = 0.7) +
      labs(title = paste("Finish Times for", input$runner_type, "Runners"),
           x = "Finish Time (hours)", y = "Count") +
      theme_minimal()
  })
}

# Run the Shiny app
shinyApp(ui = ui, server = server)

```

Shiny allows us to build interactive features in R Markdown, providing real-time feedback and making our analysis more dynamic.